Tree height and hydraulic traits shape growth responses across droughts in a temperate broadleaf forest

### 2020-07-13 manuscript/climate_sensitivity_draft.Rmd
update methods & results

We then determined the best full models for predicting $Rt$ (or $Rt_{ARIMA}$). 
To do so, we compared models with all possible combinations of candidate variables and identified the full set of models within $\Delta$AICc=1 of the very top model (that with lowest AICc). 
When a variable appeared in all top models and the sign of the coefficient was consistent across models, we viewed this as support for the acceptance/rejection of the associated prediction by the full models. 
If the variable appeared in only some of the models, we considered this partial support/rejection.

@@

Responses varied across species and by drought (Fig. **2**). 
Averaged across all droughts, $Rt$ was lowest in *Liriodendron tulipifera* (mean $Rt$ = **0.66**) and highest in *Fagus grandifolia* (mean $Rt$ = **0.99**).

@@

Taller trees showed stronger growth reductions during drought when evaluating the three drought years together and for 1966 individually, although $ln[H]$ was not significant during 1977 or 1999 individually (Tables 1, 4). When combined with other predictor variables in the full models, $ln[H]$ appeared, with negative coefficient, in all full models for the three droughts combined, in the 1966 model, and in one of the two models for 1999 (Tables 1, 5).

Crown position varied as expected with height (dominant > co-dominant > intermediate > suppressed), but with substantial variation (Fig. **3d**). When considered alone, crown position had a significant response only in the 1966 drought, during which trees with dominant crown position had the lowest $Rt$. Crown position was a much poorer predictor of $Rt$ than was height in the single-variable tests (Table 4), lending little overall support to the hypothesis that crown exposure reduces $Rt$ (Table 1).  When height was included in the model, crown position was a significant predictor in the 1999 drought, with lowest $Rt$ for suppressed and then intermediate trees. Crown position was included in some of the full models (Table 5). In 1977, where height was not included in the full model, dominant trees had the lowest $Rt$, and suppressed trees the highest. In contrast, in full models including both height and crown position (all droughts and 1999), the lowest $Rt$ was in suppressed, followed by intermediate, trees.

@@

$Rt$ was negatively correlated with $ln[TWI]$ (Tables 4-5), rejecting the idea that trees in moist microsites would be less affected by drought. Nevertheless, we tested for a negative $ln[H] *ln[TWI]$ interaction, which could indicate that smaller trees (with smaller rooting volume) are more susceptible to drought in drier microenvironments with a deeper water table. This hypothesis was rejected as the $ln[H] *ln[TWI]$ interaction was never significant (Table 4).

@@

Wood density, $LMA$, and xylem porosity were all poor predictors of $Rt$ (Tables 1,4,5). 
Wood density and $LMA$ were never significantly associated with $Rt$ in the single-variable tests (Table 4) and were therefore excluded from the full models. 
Xylem porosity was also excluded from the full models, as it had no significant influence for all droughts combined and had contrasting effects in the individual droughts: whereas ring-porous species had higher $Rt$ than diffuse- and semi-ring- porous species in the 1966 and 1999 droughts, they had lower $Rt$ in 1977 (Table **4**). 
It is noteworthy that the two diffuse-porous species in our study, *Liriodendron tulipifera* and *Fagus grandifolia*, were at opposite ends of the $Rt$ spectrum (Fig. **2**), further refuting the idea that xylem porosity is a useful predictor of $Rt$.

In contrast, $PLA_{dry}$, and $\pi_{tlp}$ were linked to drought responses (Tables 1,4,5). 
Both had consistent signs across all droughts and explained modest amounts of variation ($\Delta$AICc > 1.0) during at least one of the three droughts (Table 4), qualifying them as candidate variables for the full model. 
$PLA_{dry}$ was consistently included, with negative coefficient, in full models for the three droughts combined and for the 1966 and 1977 droughts individually (Table 5). 
$\pi_{tlp}$ was included in the top full model for all droughts combined and for the 1977 and 1999 droughts individually (Table 5).

>>>>

We then determined the best full models for predicting $Rt$ (or $Rt_{ARIMA}$). 
To do so, we compared models with all possible combinations of candidate variables and identified the full set of models within $\Delta$AICc=2 of the very top model (that with lowest AICc). 
When a variable appeared in all of these models and the sign of the coefficient was consistent across models, we viewed this as support for the acceptance/rejection of the associated prediction (Table 1). 
If the variable appeared in some but not all of these models, and its sign we was consistent across models, we considered this partial support/rejection. 
In presentation of the results below, we note instances where the $Rt_{ARIMA}$ model disagreed with the $Rt$ model, but otherwise do not discuss the $Rt_{ARIMA}$ model.

@@

Responses varied across species and by drought (Fig. 2). 
Averaged across all droughts, $Rt$ was lowest in *Liriodendron tulipifera* (mean $Rt$ = **0.66**) and highest in *Fagus grandifolia* (mean $Rt$ = **0.99**).

@@

Taller trees showed stronger growth reductions during drought when evaluating the three drought years together and for 1966 individually (Table 1; Fig. 4;). 
Specifically, $ln[H]$ appeared, with negative coefficient, in the top models and all models within $\Delta$AICc=2 of these (Tables S6-S7).
For the 1977 and 1999 droughts, $ln[H]$ did not appear in the top models, but was included, with negative coefficient, in some of the models that were statistically indistinguishable ($\Delta$AICc<2) from the top model (Tables 1, S6-S7).

Crown position varied as expected with height (dominant > co-dominant > intermediate > suppressed), but with substantial variation (Fig. **3d**). 
Crown position was never statistically significant (Table 1). 
It did, however, come out in the top models for 1977 and 1999 (Fig. 4), and was included in some of the models within $\Delta$AICc=2 of the top model for 1966 and all droughts combined (Table S6).
When $Rt_{ARIMA}$ was used as the predictor variable, crown position was never in a top model, and was included among the models within $\Delta$AICc=2 of the top model only for 1977.
Canopy position did not have a consistent influence, when included in models within $\Delta$AICc=2 of the top model, across years: dominant trees had the lowest $Rt$ (or $Rt_{ARIMA}$) in 1977, but in all other years suppressed trees had the lowest $Rt$ (Fig. 4; Tables S6-S7).

@@

$Rt$ had a signficantly negative response to $ln[TWI]$ in all drought years combined and in 1977, and a negative effect of $ln[TWI]$ was included in models within $\Delta$AICc=2 of the top model in 1999 and the $Rt_{ARIMA}$ model for 1966 (Fig. 4, Tables 4-5).
This negates the idea that trees in moist microsites would be less affected by drought. 
Nevertheless, we tested for a negative $ln[H] *ln[TWI]$ interaction, which could indicate that smaller trees (with smaller rooting volume) are more susceptible to drought in drier microenvironments with a deeper water table. 
This hypothesis was rejected, as the $ln[H] *ln[TWI]$ interaction was never the significant, and had a positive sign in any models within $\Delta$AICc=2 of the top model in which it appeared (Tables 1, S6-S7).
This term did appear, with positive coefficient, in the top $Rt_{ARIMA}$ model for all years combined (Table S7).

@@

Wood density, $LMA$, and xylem porosity were all poor predictors of $Rt$ (Tables 1,S4-S5). 
Wood density and $LMA$ were never significantly associated with $Rt$ in the single-variable tests and were therefore excluded from the full models. 
Xylem porosity was also excluded from the full models, as it had no significant influence for all droughts combined and had contrasting effects in the individual droughts: whereas ring-porous species had higher $Rt$ than diffuse- and semi-ring- porous species in the 1966 and 1999 droughts, they had lower $Rt$ in 1977 (Table S4). 
It is noteworthy that the two diffuse-porous species in our study, *Liriodendron tulipifera* and *Fagus grandifolia*, were at opposite ends of the $Rt$ spectrum (Fig. 2), further refuting the idea that xylem porosity is a useful predictor of $Rt$.

In contrast, $PLA_{dry}$, and $\pi_{tlp}$ were linked to drought responses (Fig. 4; Tables 1,S4-S7). 
Both had consistent signs across all droughts and explained modest amounts of variation ($\Delta$AICc > 1.0) during at least one of the three droughts (Table S4), qualifying them as candidate variables for the full model. 
$PLA_{dry}$ had a significant influence, with negative coefficient, in full models for the three droughts combined and for the 1966 drought individually (Fig. 4; Tables S6-S7). 
For 1977 and 1999, it was included, with negative coefficient, in some of the models within $\Delta$AICc=2 of the top model (Tables S6-S7).
$\pi_{tlp}$ was included, with negative coefficient, in the top full model for all droughts combined and for the 1977 drought individually (Fig. 4; Table 5). 
It was included in some of the models within $\Delta$AICc=2 of the top model in 1999 (Tables S6-S7).
